---
title: "cleaning_inv"
output: html_document
date: "2023-01-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load libraries
```{r}
library(tidyverse)
library(janitor)
library(tidycensus)
library(sf)
library(ggplot2)

```


#Load data
```{r}

bridges <- read_csv("inventory.csv")
data(fips_codes)

#create state lookup df
fips_states <- fips_codes%>%
  select(state_code, state_name)

#create county lookup df
fips_counties <- fips_codes%>%
  select(county, county_code)


service <- read_csv("service.csv")

service_type <- read_csv("service_type.csv")

owner <- read_csv("owner.csv")

cond <- read_csv("condition.csv")
```

#Clean data
#src = https://www.fhwa.dot.gov/bridge/mtguide.pdf
   
#Select columns, transform state codes to state names
```{r}

colnames(bridges)
library(tigris)

bridges_clean <- bridges %>%
  select(STATE_CODE_001, COUNTY_CODE_003, OWNER_022, ROUTE_PREFIX_005B, SERVICE_LEVEL_005C, ROUTE_NUMBER_005D, FEATURES_DESC_006A, FACILITY_CARRIED_007, LOCATION_009, TOLL_020, MAINTENANCE_021, YEAR_BUILT_027, TRAFFIC_LANES_ON_028A, ADT_029, SERVICE_ON_042A, DECK_COND_058, SUPERSTRUCTURE_COND_059, SUBSTRUCTURE_COND_060)
  #filter(OWNER_022 == "01")

bridges_clean <- clean_names(bridges_clean)%>%
        rename("state_code" = "state_code_001",
               "county_code" = "county_code_003")%>%
  mutate(full_code = paste(state_code,county_code))
  
bridges_clean$full_code <- str_replace_all(bridges_clean$full_code, " ", "")

#bridges_clean$long_017 <- as.double(as.character(bridges_clean$long_017))


bridges_clean <- transform(
  bridges_clean,
  state_code = replace(
    state_code,
    state_code %in% fips_states$state_code,
    setNames(fips_states$state_name, fips_states$state_code)[state_code][state_code %in% fips_states$state_code]
  )
)

fips_clean <- fips_codes%>%
  mutate(full_code = paste(state_code,county_code))%>%
  select(county, full_code)
  
fips_clean$full_code <- str_replace_all(fips_clean$full_code, " ", "")


bridges_clean <- transform(
  bridges_clean,
  full_code = replace(
    full_code,
    full_code %in% fips_clean$full_code,
    setNames(fips_clean$county, fips_clean$full_code)[full_code][full_code %in% fips_clean$full_code]
  )
)

###
bridges_clean <- transform(
  bridges_clean,
  service_level_005c = replace(
    service_level_005c,
    service_level_005c %in% service$service_level_005c,
    setNames(service$desc, service$service_level_005c)[service_level_005c][service_level_005c %in% service$service_level_005c]
  )
)

bridges_clean <- transform(
  bridges_clean,
  owner_022 = replace(
    owner_022,
    owner_022 %in% owner$owner_022,
    setNames(owner$desc, owner$owner_022)[owner_022][owner_022 %in% owner$owner_022]
  )
)

bridges_clean <- transform(
  bridges_clean,
  service_on_042a = replace(
     service_on_042a,
     service_on_042a %in% service_type$service_on_042a,
    setNames(service_type$desc, service_type$service_on_042a)[service_on_042a][service_on_042a %in% service_type$service_on_042a]
  )
)

bridges_clean <- transform(
  bridges_clean,
  deck_cond_058 = replace(
     deck_cond_058,
     deck_cond_058 %in% cond$code,
    setNames(cond$desc, cond$code)[deck_cond_058][deck_cond_058 %in% cond$code]
  )
)

bridges_clean <- transform(
  bridges_clean,
  superstructure_cond_059 = replace(
     superstructure_cond_059,
     superstructure_cond_059 %in% cond$code,
    setNames(cond$desc, cond$code)[superstructure_cond_059][superstructure_cond_059 %in% cond$code]
  )
)

bridges_clean <- transform(
  bridges_clean,
  substructure_cond_060 = replace(
     substructure_cond_060,
     substructure_cond_060 %in% cond$code,
    setNames(cond$desc, cond$code)[substructure_cond_060][substructure_cond_060 %in% cond$code]
  )
)



```




```{r}

head(bridges_clean)

bridges_selects <- bridges_clean%>%
  select(state_code, full_code, owner_022, service_on_042a, year_built_027, facility_carried_007, location_009, deck_cond_058, superstructure_cond_059, substructure_cond_060)

bridges_selects$facility_carried_007 <- str_replace_all(bridges_selects$facility_carried_007, "'", "")

bridges_selects$location_009 <- str_replace_all(bridges_selects$location_009, "'", "")


bridges_ex <- bridges_selects%>%
        rename("State" = "state_code",
               "County" = "full_code",
               "Owner" = "owner_022",
               "Service Type" = "service_on_042a",
               "Year Built" = "year_built_027",
               "Facility Carried" = "facility_carried_007",
               "Location" = "location_009",
               "Deck Condition" = "deck_cond_058",
               "Superstructure Condition" = "superstructure_cond_059",
               "Substructure Condition" = "substructure_cond_060"
               )

bridges_ex$`Deck Condition` = str_to_title(bridges_ex$`Deck Condition`)

bridges_ex$`Location` = str_to_title(bridges_ex$`Location`)

bridges_ex$`Facility Carried` = str_to_title(bridges_ex$`Facility Carried`)

bridges_ex$`Superstructure Condition` = str_to_title(bridges_ex$`Superstructure Condition`)

bridges_ex$`Substructure Condition` = str_to_title(bridges_ex$`Substructure Condition`)

head(bridges_ex)


```

```{r}

write.csv(bridges_ex, "clean_bridges.csv", row.names=FALSE)


```


#Get bad bridges

```{r}

bad <- bridges_selects %>%
    filter(deck_cond_058 == "IMMINENT FAILURE CONDITION" | superstructure_cond_059 == "IMMINENT FAILURE CONDITION" | substructure_cond_060 == "IMMINENT FAILURE CONDITION")

write.csv(bad, "bad_bridges.csv", row.names=FALSE)

```




### END ####




######Access lat/long

```{r}


1e7

bridges_cleaner <- bridges_clean %>%
  mutate(lat = lat_016/1e7)%>%
  mutate(long = long_017/1e7)


frm <- frm %>% 
   mutate(dogs = cats / 1000)

my_sf <- st_as_sf(bridges_clean, coords = c('long_017', 'lat_016'))

md_shp <- st_read("https://raw.githubusercontent.com/frankrowe/maryland-geojson/master/maryland.geojson")

md <- bridges_clean %>%
  filter(state_code == "Maryland")

md_sf <- st_as_sf(md, coords = c('long_017', 'lat_016'))

md_map_all <- 
  leaflet()%>%
  addProviderTiles(providers$CartoDB.Positron)%>%
  addPolylines(
    data=md_shp,
    weight=2,
    color="gray"
  )%>%
 addCircleMarkers(
    data=md_sf,
    color="orange",
    radius=2)

md_map_all

#plot

ggplot(my_sf) + 
  geom_sf()

```


----
  
  
  
library(maps)
library(sf)

## Get the states map, turn into sf object
US <- st_as_sf(map("state", plot = FALSE, fill = TRUE))

## Test the function using points in Wisconsin and Oregon
testPoints <- data.frame(x = c(-90, -120), y = c(44, 44))

# Make it a spatial dataframe, using the same coordinate system as the US spatial dataframe
testPoints <- st_as_sf(bridges_clean, coords = c("lat_016", "long_017"), crs = st_crs(US))

#.. and perform a spatial join!
st_join(testPoints, US)



----
  
  
  
  
library(maps)

startm <- Sys.time()
#I am doing the reverse geocoding using the function map.where from the maps package.
#Here I used three maps of the package; world, state and county
country<-map.where(database="world", 
                           usafire$longitude, usafire$latitude)
state<-map.where(database="state", 
                         usafire$longitude, usafire$latitude)
county<-map.where(database="county", 
                        bridges_clean$long_017, bridges_clean$lat_016)
endm <- Sys.time()


         ID        geometry
1 wisconsin  POINT (-90 44)
2    oregon POINT (-120 44)

library(tidygeocoder)
reverse <- bridges_clean %>%
  reverse_geocode(lat = "lat_016", long = "long_017")

rev1 <- bridges_clean %>%
  reverse_geocode(lat = "lat_016", long = "long_017", address = addr, method = "osm")

```



  my_sf <- st_set_crs(my_sf, crs = 4326)

vehicle_points <- st_as_sf(bridges_clean, 
                          coords = c("long_017", "lat_016"), 
                          crs = 4326, 
                          agr = "constant")

filter() statement with the is.na()

my_sf <- st_set_crs(bridges_clean, crs = 4326)

#Plot it:

ggplot(my_sf) + 
  geom_sf(aes(color = cluster))



```



```{r}


library(tidygeocoder)
reverse <- bridges_clean %>%
  reverse_geocode(lat = lat_016, long = long_017, method = 'osm',
                  address = address_found, full_results = TRUE)


bridges_clean %>% lookup_code(county_code)


bridges_clean <- transform(
bridges_clean,
 county_code = replace(
    county_code,
    county_code %in% fips_counties$county_code,
    setNames(fips_counties$county, fips_counties$county_code)[county_code][county_code %in% fips_counties$county_code]
  )
)

library(latlong2)
testPoints <- data.frame(x=c(-90, -120), y=c(44, 44), z=c("a", "b"))
latlong2(testPoints, to="state")

bridges_clean <- transform(
  bridges_clean,
 county_code = replace(
    county_code,
    county_code %in% fips_counties$county_code,
    setNames(fips_counties$county, fips_counties$county_code)[county_code][county_code %in% fips_counties$county_code]
  )
)



bridges_clean

library(data.table)

dt.replaceValueUsingLookup <- function(dt, col, dtLookup) {
   dt[
     dtLookup,
     on = setNames("old", col),
     (col) := new
      ]
     }

bridges_clean %>% 
   dt.replaceValueUsingLookup("state_code", fips_states)
bridges_clean

bridges_clean %>%
   gather(key = "state_code") %>%
   left_join(fips_states, by = "state_code") %>%
   spread(key = state_code, value = state_name)


bridges_clean %>%
  left_join(fips_states, by = 'state_code') %>%
  transmute(state_code = coalesce(state_name, state_code))
bridges_clean

inds <- match(bridges_clean$state_code, fips_states$state_code)
bridges_clean$state_code[!is.na(inds)] <- fips_states$state_code[na.omit(inds)]
bridges_clean


bridges_clean <- inner_join(bridges_clean, fips_states, by = c("STATE_CODE_001" = "state_code"))
bridges_clean <- inner_join(bridges_clean, fips_counties, by = c("COUNTY_CODE_003" = "county_code"))





```
```{r}
library(tidyverse)
library(sf)
library(mapview)
mapview(bridges_clean, xcol = "long_017", ycol = "lat_016", crs = 4269, grid = FALSE)
sbux_sf <- st_as_sf(bridges_clean, coords = c("long_017", "lat_016"),  crs = 4326)


```


bridges_clean <- clean_names(bridges)%>%
  
  
bridges_clean <- inner_join(bridges, fips_states, by = c("STATE_CODE_001" = "state_code"))

head(bridges_clean)



```